Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2017-10-27 TACAS2018/paper.tex
norm notation

\usepackage{microtype}
\usepackage{graphicx}
\usepackage{color}
\usepackage{amssymb}
\usepackage{amsmath}
%\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{framed}
\usepackage{listings}
\usepackage{algorithm,algpseudocode}
\usepackage{epstopdf}
\usepackage{pifont}
\usepackage{todonotes}
\usetikzlibrary{positioning, automata, shapes.arrows, calc, shapes, arrows}
\usetikzlibrary{patterns}
\usepackage{url}

@@

\item We automatically generate \emph{correct-by-construction} digital
  state-feedback observer controllers that guarantee a given safety property
  alongside stability and response time specifications.  We note that
  observers are much more complex and difficult to synthesise than standard
  full-state feedback controllers.
%
\item  Our synthesis addresses challenges created by digitalization errors
 such as quantization, and time discretization as well as delays in the
 sampling and processing times which introduce control errors in real
 implementations.
 \item we present a model for soundly evaluating reachability in a combined 
 continuous-discrete time/space. Sound tools for continuous time reachability 
 are scarce and to the authors' knowledge no tool provides a model that 
 combines both. 
%
\item We develop a model for inductive synthesis that combines CEGIS with
abstraction refinement in order to minimize the synthesis cost. In particular
we explore the idea of a multi-tiered model where different verification, 
synthesis and modelling tools/modules explore different parts of the problem
to deliver an optimal response.
%
\end{enumerate}

@@

\begin{theorem}
The error introduced by a delay $t_d$ in the control output $\vec{u}_k$ of a discrete feedback controller for a continuous LTI system can be modelled as a non-deterministic noise signal upperbound by 
\begin{align}
|\vec{y}_{k}-{\vec{y}_{t_d}}_{k}|_\infty \leq q_t, \text{ where } q_t=&(|\mat{C}(\mat{B}_{T_s}-\mat{B}_{t_s})|_1+|\mat{C}\mat{A}_{t_s}\mat{B}_{t_d}|_1)|U|_\infty
\end{align}
where $\vec{y}_{t_d}$ is the output of the system with delay $t_d$.
\end{theorem}
%
\todo{notation $| |_x$ ?}
%
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Errors generated by the numerical representation} 
\label{sec:numeric_rep}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

Our abstraction refinement CEGIS loop is similar to CEGIS with solution generalization \cite{}.
However, it provides several improvements by exploiting modularity inside the loop.
The novelty of our approach relies on the synthesis refinement phase, which enables us
to generalise an existing candidate by means of optimisation. Rather than allow unsuitable
candidates reach the verification phase, and possibly require further induction on the synthesiser,
we use an optimizing algorithm which finds a local minima for the properties of interest. 
This reduces the chances of the optimised candidate failing verification. Furthermore, it allows
us to relax the quality of the synthesis, since unlikely candidates will still be optimised during this phase. 
Optimization may also enter its own refinement loop, since it is possible for the
counterexample to indicate which properties need optimizing (i.e. the ones that are failing).

@@

by~\cite{DBLP:conf/arith/BrainTRW15}.  We examine the most relevant ones:

@@

Let $q_c=2^{-F_c},q_a=2^{-F_a},q_d=2^{-F_d}$ be the smallest difference
between two numbers in their corresponding domains.  Assuming that our
system truncates numbers to the nearest value, this means thet the error
$\delta$ introduced by the function $\mathcal{F}_{\langle * \rangle}$ is
$\delta_*=\mathcal{F}_{\langle * \rangle}(x)-x \in \left[-\frac{q_*}{2}\
\frac{q_*}{2}\right]$.  We note that in the case of floating point,the
choice of $F$ is taken from the largest multiplication found (e.g.,
$|\mat{A}|_1|\vec{x}|_\infty \mid \vec{x} \models \phi_\mathit{safety}$,
where $\phi_\mathit{safety}$ is defined in \ref{eq:safetyspec}). From these
equations we can define:
%
\begin{align*}
\nu_1 \in \left[-\frac{q_a}{2}\ \frac{q_a}{2}\right] \cap  \left[-\frac{q_c}{2}\ \frac{q_c}{2}\right],
\nu_2 \in \left[-\frac{q_d}{2}\ \frac{q_d}{2}\right]
\end{align*}

We may also bound $\nu_2'$ from \eqref{eq:uk} as follollows:
\begin{align*}
\delta\left(\mat{K}_{\langle C\rangle} \vec{x}_{\langle C\rangle} \right) &\in \left[ -p \frac{q_c}{2}\ \ p \frac{q_c}{2}\right]\\
{\vec{\nu_e}}_i &\in \left[ -(p+2) \frac{q_c}{2}\ \ (p+2) \frac{q_c}{2}\right], i \in [1, p]\\
\Rightarrow \nu_2' &\in \left[  -p(|\mat{K}_{\langle C \rangle}|_1+2) \frac{q_c}{2}\ \ p(|\mat{K}_{\langle C \rangle}|_1+2) \frac{q_c}{2}\right]
\end{align*}
Similarly we also find
\begin{align*}
\nu_1' &\in \left[  -p(|{\mat{C}_d}_{\langle C \rangle}|_1+2) \frac{q_c}{2}\ \ p(|{\mat{C}_d}_{\langle C \rangle}|_1+2) \frac{q_c}{2}\right]
\end{align*}
The proof derives from $|\mat{K}\vec{x}| \leq |\mat{K}|_1|\vec{x}|_\infty$.

%-------------------------------
\section{Stability} 
\label{appendix:stabspecification}
%-------------------------------

A discrete-time LTI system as \eqref{eq:closedloopss} is
asymptotically stable if all the roots of its characteristic
polynomial (i.e., the eigenvalues of the closed-loop matrix $A_d - B_d
K$) are inside the unity circle of the complex plane, i.e., their
absolute values are strictly less than one~\cite{astrom1997computer}
(this simple sufficient condition can be generalised, however this is
not necessary in our work).  In this paper, we express this stability
specification $\phi_\mathit{stability}$ in terms of a check known as
\emph{Jury's criterion}~\cite{fadali}: this is an easy algebraic
formula to select the entries of matrix $K$ so that the closed-loop
dynamics are shaped as desired.

@@

\begin{proof}
Let us look at equations \eqref{eq:feedback_sampled_cont} and
\eqref{eq:feedback_cont}.  When we introduce an arbitrary delay $T_d$ where
$t_d=T_d$ and $t_s=T_s-T_d$ we obtain:
%
\begin{align}
\vec{x}_{k} &=\mat{A}_{t_s}(\mat{A}_{t_d}\vec{x}_{k-1}-\mat{B}_{t_d}\mat{K}\vec{x}_{k-2})-\mat{B}\mat{K}\vec{x}_{k-1}\nonumber\\
&=  (\mat{A}_{T_s}-\mat{B}_{t_s}\mat{K})\vec{x}_{k-1}-\mat{A}_{t_s}\mat{B}_{t_d}\mat{K}\vec{x}_{k-2}
\label{eq:delay_cont}
\end{align}
%
Differentiating with the expected progression without delay, we get a delay
error:
%
\begin{align}
{\vec{x}_e}_{k}=&(\mat{B}_{T_s}-\mat{B}_{t_s})\mat{K}\vec{x}_{k-1}-\mat{A}_{t_s}\mat{B}_{t_d}\mat{K}\vec{x}_{k-2}\nonumber\\
=&(\mat{B}_{T_s}-\mat{B}_{t_s})\vec{u}_{k}-\mat{A}_{t_s}\mat{B}_{t_d}\vec{u}_{k-1}
\end{align}
Rather than explicitly calculate this error, we will find an upper bound. Let $|U|_\infty=max(|\vec{u}|_\infty \mid \vec{u} \in U)$ be the maximum infinity norm of a vector set, propagating the error to the output $\vec{y}_k$ we have:
\begin{align}
|{\vec{y}_e}_{k}|_1 \subseteq &(\mat{C}(\mat{B}_{T_s}-\mat{B}_{t_s})+\mat{C}\mat{A}_{t_s}\mat{B}_{t_d})U
\end{align}
\end{proof}
We overapproximate the effects of this vector by introducing a noise signal at the output $\nu_t=[-q_t\ q_t]$ and use the dynamics of the system without delay.

>>>>

\usepackage{microtype}
\usepackage{graphicx}
\usepackage{color}
\usepackage{amssymb}
\usepackage{amsmath}
%\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{framed}
\usepackage{listings}
\usepackage{algorithm,algpseudocode}
\usepackage{epstopdf}
\usepackage{pifont}
\usepackage{todonotes}
\usetikzlibrary{positioning, automata, shapes.arrows, calc, shapes, arrows}
\usetikzlibrary{patterns}
\usepackage{url}
\usepackage{mathtools}

\DeclarePairedDelimiter\norm{\lVert}{\rVert}

@@

\item We automatically generate \emph{correct-by-construction} digital
  state-feedback observer controllers that guarantee a given safety property
  alongside stability and response time specifications.  We note that
  observers are much more complex and difficult to synthesise than standard
  full-state feedback controllers.
%
\item  Our synthesis addresses challenges created by digitalization errors
 such as quantization, and time discretization as well as delays in the
 sampling and processing times which introduce control errors in real
 implementations.
 \item We present a model for soundly evaluating reachability in a combined 
 continuous-discrete time/space. Sound tools for continuous time reachability 
 are scarce and to the authors' knowledge no tool provides a model that 
 combines both. 
%
\item We develop a model for inductive synthesis that combines CEGIS with
abstraction refinement and synthesis optimization in order to minimize the
synthesis cost. In particular we explore the idea of a multi-tiered model where
different verification, synthesis and modelling tools/modules explore different
parts of the problem to deliver an optimal response.
%
\end{enumerate}

@@

\begin{theorem}
The error introduced by a delay $t_d$ in the control output $\vec{u}_k$ of a discrete feedback controller for a continuous LTI system can be modelled as a non-deterministic noise signal upperbound by 
\begin{align}
\norm{\vec{y}_{k}-{\vec{y}_{t_d}}_{k}}_1 \leq q_t, \text{ where } q_t=&(\norm{\mat{C}(\mat{B}_{T_s}-\mat{B}_{t_s})}_1+\norm{\mat{C}\mat{A}_{t_s}\mat{B}_{t_d}}_1)\norm{U}_1
\end{align}
where $\vec{y}_{t_d}$ is the output of the system with delay $t_d$, and $\norm{\vec{x}}_1=\sum_{i=0}^{size(\vec{x})} |\vec{x}_i|$ is the $l_1$ norm (in for sets, $\norm{X}_1 = sup \norm{\vec{x}}_1 \mid \vec{x} \in X$). 
\end{theorem}
%
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Errors generated by the numerical representation} 
\label{sec:numeric_rep}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

Our abstraction refinement CEGIS loop is similar to CEGIS with counterexample generalization~\cite{david2017program}.
However, it provides several improvements by exploiting modularity inside the loop.
The novelty of our approach relies on the synthesis refinement phase, which enables us
to generalise an existing candidate by means of optimisation. Rather than allow unsuitable
candidates reach the verification phase, and possibly require further induction on the synthesiser,
we use an optimizing algorithm which finds a local minima for the properties of interest. 
This reduces the chances of the optimised candidate failing verification. Furthermore, it allows
us to relax the quality of the synthesis, since unlikely candidates will still be optimised during this phase. 
Optimization may also enter its own refinement loop, since it is possible for the
counterexample to indicate which properties need optimizing (i.e. the ones that are failing).

@@

by~\cite{DBLP:conf/arith/BrainTRW15}.  
We examine the most relevant ones:

@@

Let $q_c=2^{-F_c},q_a=2^{-F_a},q_d=2^{-F_d}$ be the smallest difference
between two numbers in their corresponding domains.  Assuming that our
system truncates numbers to the nearest value, this means thet the error
$\delta$ introduced by the function $\mathcal{F}_{\langle * \rangle}$ is
$\delta_*=\mathcal{F}_{\langle * \rangle}(x)-x \in \left[-\frac{q_*}{2}\
\frac{q_*}{2}\right]$.  We note that in the case of floating point,the
choice of $F$ is taken from the largest multiplication found (e.g.,
$\norm{\mat{A}}_1\norm{\vec{x}}_1 \mid \vec{x} \models \phi_\mathit{safety}$,
where $\phi_\mathit{safety}$ is defined in \ref{eq:safetyspec}). From these
equations we can define:
%
\begin{align*}
\nu_1 \in \left[-\frac{q_a}{2}\ \frac{q_a}{2}\right] \cap  \left[-\frac{q_c}{2}\ \frac{q_c}{2}\right],
\nu_2 \in \left[-\frac{q_d}{2}\ \frac{q_d}{2}\right]
\end{align*}

We may also bound $\nu_2'$ from \eqref{eq:uk} as follollows:
\begin{align*}
\delta\left(\mat{K}_{\langle C\rangle} \vec{x}_{\langle C\rangle} \right) &\in \left[ -p \frac{q_c}{2}\ \ p \frac{q_c}{2}\right]\\
{\vec{\nu_e}}_i &\in \left[ -(p+2) \frac{q_c}{2}\ \ (p+2) \frac{q_c}{2}\right], i \in [1, p]\\
\Rightarrow \nu_2' &\in \left[  -p(|\mat{K}_{\langle C \rangle}|_1+2) \frac{q_c}{2}\ \ p(|\mat{K}_{\langle C \rangle}|_1+2) \frac{q_c}{2}\right]
\end{align*}
Similarly we also find
\begin{align*}
\nu_1' &\in \left[  -p(|{\mat{C}_d}_{\langle C \rangle}|_1+2) \frac{q_c}{2}\ \ p(|{\mat{C}_d}_{\langle C \rangle}|_1+2) \frac{q_c}{2}\right]
\end{align*}
The proof derives from $\norm{\mat{K}\vec{x}}_1 \leq \norm{\mat{K}}_1\norm{\vec{x}}_1$.

%%-------------------------------
%\section{Stability} 
%\label{appendix:stabspecification}
%%-------------------------------
%
%There are a number of algorithms that can be
%used for stability analysis~\cite{daes20161,Bessa16}.
%
%A discrete-time LTI system as \eqref{eq:closedloopss} is
%asymptotically stable if all the roots of its characteristic
%polynomial (i.e., the eigenvalues of the closed-loop matrix $A_d - B_d
%K$) are inside the unity circle of the complex plane, i.e., their
%absolute values are strictly less than one~\cite{astrom1997computer}
%(this simple sufficient condition can be generalised, however this is
%not necessary in our work).  In this paper, we express this stability
%specification $\phi_\mathit{stability}$ in terms of a check known as
%\emph{Jury's criterion}~\cite{fadali}: this is an easy algebraic
%formula to select the entries of matrix $K$ so that the closed-loop
%dynamics are shaped as desired.
%
%
%
%Since the canonical forms used in this work use the coefficients of the
%characteristic polynomial $S(z)$ as opposed to its roots (which are
%extremely hard to calculate for a SAT solver), we choose Jury's
%criterion~\cite{astrom1997computer} to check the stability in the
%$z$-domain in view of its efficiency.
%
%We~consider the following form for $S(z)$:
%%
%\begin{equation*}
%S(z) = z^{p}+a_1z^{p-1}+\cdots+a_{p-1}z+a_p.
%\end{equation*}
%
%and build the matrix
%$\mat{M} \in \mathbb{R}^{p \times p}$ from its coefficients such that:
%%
%\begin{align*}
%\mat{M}_{1,j}&=a_j\\
%\mat{M}_{i+1,j}&=\left\{
%\begin{array}{ll}
%0,&j>p'\\
%\mat{M}_{i,j}-\mat{M}_{i,p'-j+1}\frac{\mat{M}_{i,1}}{\mat{M}_{i,p'}} &j\leq p'
%\end{array}
%\right.\\
%p'&=p-i+1
%\end{align*}
%%
%$S(z)$ is the characteristic polynomial of a stable system if and
%only if the following conditions hold~\cite{astrom1997computer}:
%\begin{align*}
%\phi_\mathit{stability}(S) \equiv
%(S(1) > 0) \wedge ((-1)^p S(-1) > 0) \wedge (|a_p| < 1) \wedge \bigwedge\limits_{i=1}^p (\mat{M}_{i,1} > 0).
%\end{align*}
%

@@

\begin{proof}
Let us look at equations \eqref{eq:feedback_sampled_cont} and
\eqref{eq:feedback_cont}.  When we introduce an arbitrary delay $T_d$ where
$t_d=T_d$ and $t_s=T_s-T_d$ we obtain:
%
\begin{align}
\vec{x}_{k} &=\mat{A}_{t_s}(\mat{A}_{t_d}\vec{x}_{k-1}-\mat{B}_{t_d}\mat{K}\vec{x}_{k-2})-\mat{B}\mat{K}\vec{x}_{k-1}\nonumber\\
&=  (\mat{A}_{T_s}-\mat{B}_{t_s}\mat{K})\vec{x}_{k-1}-\mat{A}_{t_s}\mat{B}_{t_d}\mat{K}\vec{x}_{k-2}
\label{eq:delay_cont}
\end{align}
%
Differentiating with the expected progression without delay, we get a delay
error:
%
\begin{align}
{\vec{x}_e}_{k}=&(\mat{B}_{T_s}-\mat{B}_{t_s})\mat{K}\vec{x}_{k-1}-\mat{A}_{t_s}\mat{B}_{t_d}\mat{K}\vec{x}_{k-2}\nonumber\\
=&(\mat{B}_{T_s}-\mat{B}_{t_s})\vec{u}_{k}-\mat{A}_{t_s}\mat{B}_{t_d}\vec{u}_{k-1}
\end{align}
Rather than explicitly calculate this error, we will find an upper bound. Let $U = \bigcup \vec{u}_k$. Propagating the error to the output $\vec{y}_k$ we have:
\begin{align}
{\vec{y}_e}_{k} \subseteq &\mat{C}(\mat{B}_{T_s}-\mat{B}_{t_s})U\oplus\mat{C}\mat{A}_{t_s}\mat{B}_{t_d}U\\
\Rightarrow \norm{{\vec{y}_e}_{k}}_1 \leq & (\norm{\mat{C}(\mat{B}_{T_s}-\mat{B}_{t_s})}_1+\norm{\mat{C}\mat{A}_{t_s}\mat{B}_{t_d}}_1)\norm{U}_1
\end{align}
\end{proof}
We overapproximate the effects of this vector by introducing a noise signal at the output $\nu_t=[-q_t\ q_t]$ and use the dynamics of the system without delay.
